---
title: Local Level Model
author: admin
date: '2021-07-01'
summary: 'In this short project, we present some exercices regarding the Local Level Model. We apply the Kalman filter to simulated data and illustarte the pile-up problem that may occur for maximum likelihood estimation and small sample size.'
slug: []
categories:
#  - Macroeconometrics
#  - Local Level Model
#  - Kalman Filter
tags:
  - Macroeconometrics
  - Local Level Model
  - Kalman Filter
links:
- icon: twitter
  icon_pack: fab
  name: Follow
  url: https://twitter.com/georgecushen
slides: ""
math: true
description: Description for the page
bibliography: [bibliography.bib]
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, fig.height = 4, warning = FALSE, out.width = "100%"
)
library(tidyverse)
library(lubridate)
library(zoo)
library(plotly)
library(here)
library(tidycensus)
library(rvest)
library(knitr)
library(kableExtra)
library(readxl)
library(htmltools)
library(MTS)
```


```{r}
# Move this css tag outside the chunk to control the width of text
# on the page.
# <style type="text/css">
# .main-container {
#   max-width: 1000px;
#   margin-left: auto;
#   margin-right: auto;
# }
# </style>
```

# Framework

In the exercise, I was provided with monthly export data of a Swiss industry. The aim was to forecast exports for several different regions, including Europe, the United States as well as total exports. In this short project description, I will focus on the procedure to forecast total exports of the Swiss industry.


Below, the figure shows the original series of total exports of this Swiss industry. The series shows strong seasonality and a positive trend over time. The series starts in 1993 and total exports peak in october 2014 at 2.28 billion CHF.


```{r, include = FALSE}

# clear variables:
rm(list=ls(all=TRUE))

```

test
